---
title: "Assignment 6 - Split-plot randomization"
author: "Micah Jones"  
format:
  html:
    code-fold: true 
    embed-resources: true
    toc: true
    number-sections: true
    theme: cerulean
---

# Introduction  
This assignment will ask you to conduct a split-plot RCBD randomization. 

For that, you will first need to watch (and code along with) the following split-plot randomization video:

- split-plot randomization: https://youtu.be/rXTp3Vw_770?si=MajBK1aZ221orfpa&t=3401 (starts at 00:56:41, watch till end, about 50-min total)

- The partial script for the video above is in our GitHub repository, 04-classcode, **02-13_spp_randomization_partial.qmd**.


**AFTER watching the video and developing the code in class**, then start working on this HW assignment.  

The different sections representing what we did in class were left in the script to provide you a scaffold to start from.  

# Instructions  
  - Move this script into the `code` subfolder of your `05_spp` project folder  
  - On this script, fill your name under the `title` section of the YAML  
  
# Goal of this assignment  
The goal is this assignment is for you to **design a split-plot RCBD design applicable to your area of study**.  

This could be from your own experience, an example from the literature, or something you come up entirely on your own (as long as it makes sense).  

In the headings below, you will answer a couple of questions related to study objectives and treatment design, and then move on to randomize your study, create a plot layout, and export the designed table to file.  

# Setup  
Here is where we load the packages we will use.  
Make sure to load each package in the chunk below.  
```{r setup, warning=FALSE, message=FALSE}
library(tidyverse)
```



# Treatment design  
## Objectives  

  I will be completing this assignment as a split-plot version of a study I am currently working on, as that design would also make sense for it (the actual study is an RCBD design). The objectives are to determine the effects of repeated application of a microbial inoculant called LEM (Local Effective Microorganisms) to an intensive vegetable cropping system, and of three different winter crops, on crop yield, marketability, and soil nutrition.


## Treatment design explanation   
  
  The factors of interest are the rate of LEM application, at the levels of 3.6 L/m^2, 1.8 L/m^, and 0 L/m^2, and the winter crop (kale, carrots, or forage crimson clover). For the purposes of this assignment, I will be treating winter crop as the whole-plot treatment factor, and the LEM rate as the split-plot treatment factor. This experiment has 4 blocks, so there are 12 whole plots and 36 split plots.


## Code  
Here develop the code to setup your design (like in class).   



# Split-plot in RCBD  
## WP randomization  
```{r wp_rcbd_rand}
set.seed(123)

wp_rcbd_rand <- crossing(winter_crop = c("Kale", "Carrot", "Clover"),
                         block = 1:4
                         ) %>%
  arrange(block) %>%
  group_by(block) %>%
  mutate(winter_crop = sample(winter_crop)) %>%
  mutate(wplot = (block*100) + 1:3) %>% 
  ungroup()

wp_rcbd_rand
```


## SP randomization  
```{r sp_rcbd_rand}
set.seed(123)

sp_rcbd_rand  <- crossing(lemrate_lm2 = c(0, 1.8, 3.6),
                          wplot = wp_rcbd_rand$wplot) %>%
  arrange(wplot) %>%
    group_by(wplot) %>%
  mutate(lemrate_lm2 = sample(lemrate_lm2)) %>%
  mutate(splot = paste0(wplot, "-", 1:3)) %>%
  ungroup()


sp_rcbd_rand
```

## Joining WP and SP into one object
```{r spp_rcbd_rand}
spp_rcbd_rand <- wp_rcbd_rand %>%
  right_join(sp_rcbd_rand) %>%
  mutate(trtname = paste0(winter_crop, "+", lemrate_lm2))

spp_rcbd_rand

```


## Plot layout  
```{r spp_rcbd_layout, warning=FALSE}
spp_rcbd_layout <- spp_rcbd_rand %>%
  # Defining number of rows and columns in the field for plotting purposes
  mutate(row = rep(1:4, each = 9)) %>%
  mutate(col = rep(seq(from = 1, 
                       to = 9, #total number of treatments, 
                       by = 1), 
                   times = 4 #total number of blocks
  )) %>%
  # Creating plot and trtname labels for plotting
  mutate(lab = paste0(splot, "\n", trtname))

spp_rcbd_layout
```


```{r spp rcbd map, warning=FALSE}
ggplot(data = spp_rcbd_layout, aes(x = row, 
                                   y = -col)) +
    geom_tile(aes(fill = factor(lemrate_lm2)),
            color = "black",
            show.legend = F) +
  geom_tile(aes(color = winter_crop),
            show.legend = F,
            size = 2,
            linetype = 2,
            fill = NA
            ) +

    scale_fill_manual(values = c("#e5f0db",
                               "#b1d095",
                              "#5e813f")) +

  geom_text(aes(label = splot), fontface = "bold", vjust = -.5)+
  geom_text(aes(label = trtname), vjust = 1)+
  labs(title = "Split-plot RCBD randomization of a\n3 Winter Crop (whole-plot) x 3 LEM rate (split-plot) treatment design") +
  theme_void()
```

# Exporting to file  
```{r}
spp_rcbd_rand %>%
  # Rearranging columns  
  dplyr::select(block, wplot, winter_crop, 
                splot, lemrate_lm2,
                trtname) %>%
  write_csv("../output/spp_rcbd_randomization_table_lem.csv")


```
```{r}
ggsave("../output/spp_rcbd_layoutmap_lem.png",
       width = 6,
       height = 5,
       bg = "white")
```


# Submitting your work  
Once you have developed all the code and answers, make sure to Render this quarto file.  

**Notes on rendering**:  

- Make sure to render your work and inspect how the final html look like.  
- If it does not look professional for whatever reason, then fix the issue, re-render it, recheck.  
- Only send me your work once your html file looks professional.  
  - **DO NOT** delete the file's heading levels (# and ##). They set up the proper heading 1 and 2 levels, and I use them to guide my grading.  
  - If a given chunk is also outputting warnings or messages, inhibit this behavior by changing the chunk options `message` and `warning` to `FALSE` (I'll demo this in a moment).  
  
  - If, after rendered, 2 lines of text are connected and you wish to "break line" between them, add 2 extra spaces after the first one (I'lld demo this in a moment).  

After rendering, an .html file will be created on your `code` folder.  

Rename this file to `Assignment-06-spp-LASTNAME.html`.    

For ex., mine would be `Assignment-06-spp-Bastos.html`.

Submit your work by uploading the **html** file to **eLC Assigment #6 - Split-plot randomization** by Feb 20 11:59 pm.  
  




